---
title: "AIM abundance-impact"
author: "Helen Sofaer"
date: "January 7, 2020"
output: 
  html_document:
  fig_width: 4
  fig_height: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(psych)
library(lme4)
library(effects)
theme_set(theme_bw())
statebound = map_data('state')

setwd("C:/Users/hsofaer/Documents/HelenProjects/AI_AIM")
AIM.PlotDiv <- read_rds("AIM_PlotDiv_withCovariates_2Jan2020.rds")

## reshape to wide to look at exotic cover vs native community attributes
AIM.wide <- AIM.PlotDiv %>%
  # first have to gather:
  pivot_longer(c(TotalPctCover, Richness, DiversityShannon, DiversityInvSimpson,
                 EvennessEvar, EvennessPielou),
               names_to = c("variable"), values_to = "value") %>%
  unite(variable_Exotic, variable, Exotic, sep = "_") %>%
  pivot_wider(names_from = variable_Exotic, values_from = value) %>%
  mutate(TotalPctCover_AllExceptI = TotalPctCover_I + AbsUnkCover,
         TotalCover = TotalPctCover_NA + TotalPctCover_NI + TotalPctCover_I + TotalPctCover_N,
         RelCov_I = TotalPctCover_I / TotalCover)

```

## AIM: basics 
### Still have issue of cover from unknown species; treatment of NI species cover

```{r}

##  Histogram of non-native cover:
AIM.wide %>%
  ggplot() +
  geom_histogram(aes(TotalPctCover_I)) +
  xlab('Total non-native percent cover')

AIM.wide %>%
  dplyr::select(Plot, TotalCover, 
                TotalPctCover_NA, TotalPctCover_NI, TotalPctCover_I, TotalPctCover_N,
                RelCov_I) %>%
  mutate(RelCov_N = TotalPctCover_N / TotalCover,
         RelCov_NA = TotalPctCover_NA / TotalCover,
         RelCov_NI = TotalPctCover_NI / TotalCover) %>%
  dplyr::select(-contains('TotalPctCover')) %>%
  arrange(-RelCov_I) %>%
  mutate(PlotRank_RelInv = rank(desc(RelCov_I), ties.method = 'first')) %>%
  pivot_longer(contains('RelCov_'), names_to = 'SpType', values_to = 'RelativeCover') %>%
  mutate(SpType = gsub("RelCov_", "", SpType)) %>%
  ggplot(aes(x = PlotRank_RelInv, y = RelativeCover)) +
  geom_bar(aes(fill = SpType), 
           stat = 'identity',
           position = 'stack') +
  scale_fill_viridis_d() +
  xlab('Rank of plot from more to less relative non-native cover')

# AIM.wide %>%
#   filter(RelUnkCover > .05) %>%
#   nrow()

# Absolute and relative cover of unk status sp
AIM.wide %>%
  ggplot(aes(AbsUnkCover, RelUnkCover * 100)) +
  geom_point(alpha = .2) +
  xlab('Absolute cover unk status species (NA + NI)') +
  ylab('Relative cover unk status species (NA + NI)') +
  ggtitle('For now, excluding plots where relative cover > 5% (~3500 of ~15.5K)') +
  geom_hline(yintercept = 5)

# plot absolute vs relative invasive cover:
AIM.wide %>%
  ggplot(aes(TotalPctCover_I, RelCov_I * 100)) +
  geom_point(alpha = .2) +
  geom_abline(intercept = 0, slope = 1) +
  ggtitle('Focusing on absolute cover but could consider relative for some responses')
```

``` {r, fig.width = 12, fig.height = 18}
##  Spatial plots: absolute inv cover, relative inv cover, unk status (> .5%)
AIM.wide %>%
  dplyr::select(Plot, Long, Lat, 
                TotalPctCover_I, RelCov_I, RelUnkCover) %>%
  mutate(HighUnk = factor(ifelse(RelUnkCover > 0.05, 1, 0)),
         # put on percentage scale to match total
         RelCov_I = RelCov_I * 100,
         RelUnkCover = RelUnkCover * 100) %>%
  pivot_longer(contains('Cov'), names_to = 'variable', values_to = 'value') %>%
  ggplot() +
  geom_point(aes(Long, Lat, color = value, shape = HighUnk),
             size = 1, alpha = .5) +
  scale_color_viridis_c() + 
  facet_wrap(~ variable, ncol = 1) +
  geom_path(aes(long, lat, group = group), 
            data = statebound[statebound$long < -100, ]) +
  theme_void() +
  coord_map("albers", lat0 = 23, lat1 = 29.5) 
```

## Raw relationships between non-native abundance & community metrics
```{r}
##  Exotic cover vs native metrics:
AIM.wide %>%
  filter(RelUnkCover < .05) %>%
  pivot_longer(ends_with('_N'), names_to = "NativeMetric", values_to = "value") %>%
  ggplot(aes(TotalPctCover_I, value)) +
  geom_point(alpha = .2) +
  facet_wrap(~ NativeMetric, scales = 'free') +
  xlab('Total non-native percent cover') +
  ylab('Value of native community metric') +
  geom_smooth(se = FALSE)

```

### Relative unk cover vs. potential responses of interest
####  Before excluding plots with rel cover > 5%, did see some relationships 

```{r}
AIM.wide %>%
  filter(RelUnkCover < .05) %>%
  select(Plot, RelUnkCover, TotalPctCover_I, DiversityInvSimpson_N, EvennessEvar_N, Richness_N) %>%
  pivot_longer(c(TotalPctCover_I, DiversityInvSimpson_N, EvennessEvar_N, Richness_N), 
               names_to = "variable", values_to = 'value') %>%
  ggplot(aes(RelUnkCover, value)) +
  geom_point(alpha = .2) +
  facet_wrap(~ variable, scales = 'free') +
  geom_smooth(se = FALSE) +
  ggtitle('Unk. status below 5% seems ok')
```

## Non-native vs native richness & diversity 
### correlations across space often positive b/c of env heterogeneity...not much of pattern here
```{r}
AIM.wide %>%
  filter(RelUnkCover < .05) %>%
  ggplot(aes(Richness_I, Richness_N)) +
  geom_point(alpha = .2) +
  xlab("Non-native richness") +
  ylab('Native richness') +
  geom_smooth(se = FALSE)

AIM.wide %>%
  filter(RelUnkCover < .05) %>%
  ggplot(aes(DiversityInvSimpson_I, DiversityInvSimpson_N)) +
  geom_point(alpha = .2) +
  xlab("Non-native diversity (Inverse Simpson)") +
  ylab('Native diversity (Inverse Simpson)') +
  geom_smooth(se = FALSE)

AIM.wide %>%
  filter(RelUnkCover < .05) %>%
  ggplot(aes(EvennessEvar_I, EvennessEvar_N)) +
  geom_point(alpha = .2) +
  xlab("Non-native evenness") +
  ylab('Native evenness') +
  geom_smooth(se = FALSE)
```

## Non-native cover versus metrics of disturbance 
### Focus on fire history and human modification index
### Fire is really important; disturbance index not so much
```{r}
AIM.wide %>%
  filter(RelUnkCover < .05) %>%
  ggplot(aes(factor(Burned), TotalPctCover_I)) +
  geom_boxplot()

AIM.wide %>%
  filter(RelUnkCover < .05) %>%
  ggplot(aes(gHumanMod, TotalPctCover_I)) +
  geom_point(alpha = .2)

```

##  Can native community attributes be captured by ecoregion?
```{r}
AIM.wide %>%
  filter(RelUnkCover < .05) %>%
  ggplot(aes(L3_KEY, DiversityInvSimpson_N)) +
  geom_boxplot() +
  xlab('L3 EPA ecoregion') +
  theme(axis.text.x = element_blank())

AIM.wide %>%
  filter(RelUnkCover < .05) %>%
  ggplot(aes(L3_KEY, DiversityShannon_N)) +
  geom_boxplot() +
  xlab('L3 EPA ecoregion') +
  theme(axis.text.x = element_blank())

AIM.wide %>%
  filter(RelUnkCover < .05) %>%
  ggplot(aes(L3_KEY, Richness_N)) +
  geom_boxplot() +
  xlab('L3 EPA ecoregion') +
  theme(axis.text.x = element_blank())

AIM.wide %>%
  filter(RelUnkCover < .05) %>%
  mutate(Icover_bin = cut(TotalPctCover_I, breaks = c(0, 1, 5, 10, 20, 50, max(TotalPctCover_I)))) %>%
  ggplot(aes(L4_KEY, DiversityInvSimpson_N)) +
  geom_point(aes(color = Icover_bin), alpha = .4) +
  xlab('L4 EPA ecoregion') +
  theme(axis.text.x = element_blank()) +
  scale_color_viridis_d(name = 'I cover bin') +
  ggtitle('Higher I cover where N diversity is low for ecoregion?')

AIM.wide %>%
  filter(RelUnkCover < .05) %>%
  mutate(Icover_bin = cut(TotalPctCover_I, breaks = c(0, 1, 5, 10, 20, 50, max(TotalPctCover_I)))) %>%
  ggplot(aes(L4_KEY, Richness_N)) +
  geom_point(aes(color = Icover_bin), alpha = .4) +
  xlab('L4 EPA ecoregion') +
  theme(axis.text.x = element_blank()) +
  scale_color_viridis_d(name = 'I cover bin') +
  ggtitle('Higher I cover where N richness is low for ecoregion?')
```

##  Distribution of potential response variables:
```{r}
AIM.wide %>%
  filter(RelUnkCover < .05) %>%
  pivot_longer(ends_with('_N'), names_to = 'variable', values_to = 'value') %>%
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~ variable, scales = 'free')

pairs.panels(AIM.wide[, grepl('_N$', names(AIM.wide))])
```

##  Abundance-impact in burned vs unburned plots:
```{r}
AIM.wide %>%
  filter(RelUnkCover < .05) %>%
  mutate(Burned = factor(Burned,
                         levels = c('0', '1'),
                         labels = c('unburned', "burned"))) %>%
  dplyr::select(Plot, Burned,
                TotalPctCover_I, RelCov_I,
                TotalPctCover_N, Richness_N, 
                DiversityShannon_N, DiversityInvSimpson_N, EvennessEvar_N) %>%
  pivot_longer(ends_with('_N'), names_to = 'Nvariable', values_to = 'value') %>%
  ggplot(aes(TotalPctCover_I, value)) +
  geom_point() +
  facet_wrap(Nvariable ~ Burned, scales = 'free') +
  geom_smooth(se = FALSE)
```

##  Example lmer: can ecoregion capture underlying variation in native community, revealing invasion impacts?
```{r}
AIM.use = AIM.wide %>%
  filter(RelUnkCover < .05) %>%
  mutate(Burned = factor(Burned))

m.Nrich <- lmer(Richness_N ~ TotalPctCover_I*Burned + gHumanMod + (1|L4_KEY),
                data = AIM.use)
summary(m.Nrich)
plot(predictorEffect("TotalPctCover_I", m.Nrich))
```
