---
title: "CTAP_code_datapaper"
author: "Lais Petri"
date: "10/21/2021"
output: html_document
notes: David Zaya sent Laís the data via email on Sep 30 2021.
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages
```{r}
library(tidyverse)
library(lubridate)
```

Paths to data
```{r}
shared_data_dir <- "/home/shares/neon-inv/data_paper"
data_raw <- file.path(shared_data_dir, "data_by_dataset/archived")
data_output <- file.path(shared_data_dir,"data_by_dataset")
data_extra <- file.path(shared_data_dir, "code_by_dataset/extra_csv_CTAP")

```

Importing site information
```{r}
site_info <- read.csv(file.path(data_extra, "tableOfSiteVisitsFinal.csv"), header = TRUE)
glimpse(site_info)

```

Importing dataset information
```{r}
load(file.path(data_raw, "rawVegData2.RData"))
##file names: g, gshr, gtree, w, wshr, wtree, f, fshr, ftree
glimpse(g)
glimpse(gshr)
glimpse(gtree)
glimpse(w)
glimpse(wshr)
glimpse(wtree)
glimpse(f)
glimpse(fshr)
glimpse(ftree)

```

Organizing Wetlands data

```{r}
# files w, wshr, wtree

wet <- w %>% 
  select(SiteID, Genus, Species, PYear, PMonth, 35:54) %>% 
  gather(Median1:Median20, key = "Class", value = "PctCov") %>% 
  mutate(PctCov = ifelse(is.na(PctCov), 0, PctCov),
        OriginalSpeciesName =paste0(Genus, " ", Species),
        #removes any space after the end of any column containing characters
        across(where(is.character), str_trim)) %>% 
  select(-Genus, -Species, -Class) %>% 
  filter(OriginalSpeciesName != "None") %>% 
  group_by(SiteID, PYear, PMonth, OriginalSpeciesName) %>% 
  mutate(SampledAreaOriginal = .25, 
         # PctCov = mean(PctCov)/SampledArea, # Ian contacted me and this doesn't agree with the rest of the datasets
         PctCov = mean(PctCov), 
         code = 0, 
         layer = "herbaceous") %>% 
  distinct() %>% 
  ungroup() %>% 
  filter(PctCov > 0)

wettree <- wtree %>% 
  select(SiteID, PYear, PMonth, Squmetr, Genus, Species, SzclassM, PCount) %>% 
  filter(!is.na(SzclassM),
         SzclassM!=0) %>% 
  mutate(OriginalSpeciesName =paste0(Genus, " ", Species),
        #removes any space after the end of any column containing characters
        across(where(is.character), str_trim)) %>% 
  # select(-Genus, -Species, -SzclassM, -PCount) %>% 
  filter(OriginalSpeciesName != "None") %>% 
  rename(SampledAreaOriginal = Squmetr) %>% 
  mutate(BA = ((0.00007854*(SzclassM^2))*PCount), #basal area in m²
         PctCovPartial = (BA*100)/SampledAreaOriginal, #% ;  same as (BA/SampledAreaOriginal)*100
         code = 1,
         layer = "tree") %>% 
  group_by(SiteID, PYear, PMonth, OriginalSpeciesName) %>%
  mutate(PctCov = sum(PctCovPartial)) %>% 
  select(colnames(wet)) %>% 
  distinct() %>% 
  ungroup() %>% 
  filter(PctCov > 0)

wetfinal <- wet %>% 
  rbind(wettree) 

wetMultipleMonth <- wetfinal %>% 
  select(SiteID, PYear, PMonth) %>% 
  distinct() %>% 
  group_by(SiteID, PYear) %>% 
  summarize(n_obs = n()) %>% 
  filter(n_obs > 1) %>% # yes there are 23 instances
  ungroup()

wettest <- wetfinal %>% 
  group_by(SiteID, PYear) %>% 
  count(PMonth) %>% ungroup() # count the obaservations per site, year, transect and month

MonthsKeep <- wetMultipleMonth %>% 
  left_join(wettest) %>% # add information on the number of observations
  group_by(SiteID, PYear) %>% 
  top_n(1, n) %>% # filter the months with larger number of observations to be the ones we keep
  group_by(SiteID, PYear) %>% 
  slice(which.max(PMonth)) %>%  # there was one instance that a month had the excat same number of observations, so keeping the latest month of data collection
  select(-n_obs, -n)

# fixing the database
wetfinal <- wetfinal %>% 
  left_join(MonthsKeep, by = c("SiteID", "PYear")) %>% 
  mutate(PMonth = ifelse(is.na(PMonth.y), PMonth.x, PMonth.y)) %>% 
  select(-PMonth.y, -PMonth.x)

wetfinal <- wetfinal %>%
  mutate(Step1=1-(PctCov/100)) %>%
  group_by(SiteID, PYear, PMonth, OriginalSpeciesName) %>% 
  #creates a column to indicate if in that particular PLOT, there is more than one strata, meaning 'ground' and 'tree' layers. If so, a 'Y' is added, even if a single species was sampled in the tree layer
  mutate(RecordedStrata = ifelse(sum(code)>=1, "Y", "N")) %>% 
  select(-code) %>% 
  ungroup() %>% 
  group_by(SiteID, PYear, PMonth, OriginalSpeciesName) %>% 
  mutate(PctCov_100=(1-(prod(Step1)))*100) %>% 
  summarise(PctCov = sum(PctCov),
            SampledAreaOriginal = max(SampledAreaOriginal),
            RecordedStrata = dplyr::first(RecordedStrata),
            PctCov_100=dplyr::first(PctCov_100),
            NumberOfStrata = n_distinct(layer)) %>% 
  ungroup() %>% 
  mutate(Transect = 1) 

 wetfinal %>% summarize(min = min(PctCov, na.rm = TRUE),
            mean = mean(PctCov, na.rm = TRUE),
            max = max(PctCov, na.rm = TRUE),
            min_100 = min(PctCov_100, na.rm = TRUE),
            mean_100 = mean(PctCov_100, na.rm = TRUE),
            max_100 = max(PctCov_100, na.rm = TRUE))
  
```

Organizing Grasslands data

```{r}
# files g, gshr, gtree

grass <- g %>% 
  select(SiteID, Genus, Species, PYear, PMonth, 35:54) %>% 
  gather(Median1:Median20, key = "Class", value = "PctCov") %>% 
  mutate(PctCov = ifelse(is.na(PctCov), 0, PctCov),
        OriginalSpeciesName =paste0(Genus, " ", Species),
        #removes any space after the end of any column containing characters
        across(where(is.character), str_trim)) %>% 
  select(-Genus, -Species, -Class) %>% 
  filter(OriginalSpeciesName != "None") %>% 
  group_by(SiteID, PYear, PMonth, OriginalSpeciesName) %>% 
  mutate(SampledAreaOriginal = .25, 
         # PctCov = mean(PctCov)/SampledArea, # Ian contacted me and this doesn't agree with the rest of the datasets
         PctCov = mean(PctCov),
         code = 0, 
         layer = "herbaceous") %>% 
  distinct() %>% 
  ungroup() %>% 
  filter(PctCov > 0)

grasstree <- gtree %>% 
  select(SiteID, PYear, PMonth, Squrmetrs, Genus, Species, SzclassM, PCount) %>% 
  filter(!is.na(SzclassM),
         SzclassM!=0) %>% 
  mutate(OriginalSpeciesName =paste0(Genus, " ", Species),
        #removes any space after the end of any column containing characters
        across(where(is.character), str_trim),
        Squrmetrs = ifelse(is.na(Squrmetrs), 2050, Squrmetrs)) %>% 
  filter(OriginalSpeciesName != "None") %>% 
  rename(SampledAreaOriginal = Squrmetrs) %>% 
  mutate(BA = ((0.00007854*(SzclassM^2))*PCount), #basal area in m²
         PctCovPartial = (BA*100)/SampledAreaOriginal, #%
         code = 1,
         layer = "tree") %>% 
  group_by(SiteID, PYear, PMonth, OriginalSpeciesName) %>%
  mutate(PctCov = sum(PctCovPartial)) %>% 
  select(colnames(grass)) %>% 
  distinct() %>% 
  ungroup() %>% 
  filter(PctCov > 0)

grassfinal <- grass %>% 
  rbind(grasstree) %>% 
  mutate(Step1=1-(PctCov/100)) %>%
  group_by(SiteID, PYear, PMonth, OriginalSpeciesName) %>% 
    #creates a column to indicate if in that particular PLOT, there is more than one strata, meaning 'ground' and 'tree' layers. If so, a 'Y' is added, even if a single species was sampled in the tree layer
  mutate(RecordedStrata = ifelse(sum(code)>=1, "Y", "N")) %>% 
  select(-code) %>% 
  ungroup() %>% 
  group_by(SiteID, PYear, PMonth, OriginalSpeciesName) %>% 
  mutate(PctCov_100=(1-(prod(Step1)))*100) %>% 
  summarise(PctCov = sum(PctCov),
            SampledAreaOriginal = max(SampledAreaOriginal),
            RecordedStrata = dplyr::first(RecordedStrata),
            PctCov_100=dplyr::first(PctCov_100),
            NumberOfStrata = n_distinct(layer)) %>% 
  ungroup() %>% 
  mutate(Transect = 1)

grassfinal %>% summarize(min = min(PctCov, na.rm = TRUE),
            mean = mean(PctCov, na.rm = TRUE),
            max = max(PctCov, na.rm = TRUE),
            min_100 = min(PctCov_100, na.rm = TRUE),
            mean_100 = mean(PctCov_100, na.rm = TRUE),
            max_100 = max(PctCov_100, na.rm = TRUE))

```

Organizing Forests data

```{r}
# files f, fshr, ftree

forest <- f %>% 
  select(SiteID, Transect, Genus, Species, PYear, PMonth, 26:35) %>% 
  gather(Median1:Median10, key = "Class", value = "PctCov") %>% 
  mutate(PctCov = ifelse(is.na(PctCov), 0, PctCov),
        OriginalSpeciesName =paste0(Genus, " ", Species),
        #removes any space after the end of any column containing characters
        across(where(is.character), str_trim),
        PYear = ifelse(SiteID=="049601F" & Transect==1 & PYear==2020 |
                          SiteID=="118502F" & Transect==2 & PYear==2020, 2002, PYear)) %>% 
  select(-Genus, -Species, -Class) %>% 
  filter(OriginalSpeciesName != "None") %>% 
  group_by(SiteID, PYear, PMonth, Transect, OriginalSpeciesName) %>% 
  mutate(SampledAreaOriginal = .25, 
         PctCov = mean(PctCov), 
         code = 0, 
         layer = "herbaceous") %>% 
  distinct() %>% 
  ungroup() %>% 
  filter(PctCov > 0)

fortree <- ftree %>% 
  select(SiteID, PYear, PMonth, Transect, Genus, Species, SzclassM, PCount) %>% 
  filter(!is.na(SzclassM),
         SzclassM!=0) %>% 
  mutate(OriginalSpeciesName =paste0(Genus, " ", Species),
        #removes any space after the end of any column containing characters
        across(where(is.character), str_trim),
         SampledAreaOriginal = 500,
        PYear = ifelse(SiteID=="049601F" & Transect==1 & PYear==2020 |
                          SiteID=="118502F" & Transect==2 & PYear==2020, 2002, PYear)) %>% 
  # select(-Genus, -Species, -SzclassM, -PCount) %>% 
  filter(OriginalSpeciesName != "None") %>% 
  mutate(BA = ((0.00007854*(SzclassM^2))*PCount), #basal area in m²
         PctCovPartial = (BA*100)/SampledAreaOriginal, #%
         code = 1,
         layer = "tree") %>% 
  group_by(SiteID, PYear, PMonth, Transect, OriginalSpeciesName) %>%
  mutate(PctCov = sum(PctCovPartial)) %>% 
  select(colnames(forest)) %>% 
  distinct() %>% 
  ungroup() %>% 
  filter(PctCov > 0)

forestfinal <- forest %>% 
  rbind(fortree) 
  
forestMultipleMonth <- forestfinal %>% 
  select(SiteID, Transect, PYear, PMonth) %>% 
  distinct() %>% 
  group_by(SiteID, Transect, PYear) %>% 
  summarize(n_obs = n()) %>% 
  filter(n_obs > 1) %>% # yes there are 23 instances
  ungroup()
  
foresttest <- forestfinal %>% 
  group_by(SiteID, PYear, Transect) %>% 
  count(PMonth) %>% ungroup() # count the obaservations per site, year, transect and month

MonthsKeep <- forestMultipleMonth %>% 
  left_join(foresttest) %>% # add information on the number of observations
  group_by(SiteID, PYear, Transect) %>% 
  top_n(1, n) %>% # filter the months with larger number of observations to be the ones we keep
  group_by(SiteID, PYear, Transect) %>% 
  slice(which.max(PMonth)) %>%  # there was one instance that a month had the excat same number of observations, so keeping the latest month of data collection
  select(-n_obs, -n)

# fixing the database
forestfinal <- forestfinal %>% 
  left_join(MonthsKeep, by = c("SiteID", "PYear", "Transect")) %>% 
  mutate(PMonth = ifelse(is.na(PMonth.y), PMonth.x, PMonth.y)) %>% 
  select(-PMonth.y, -PMonth.x)
  
forestfinal <- forestfinal %>% 
  mutate(Step1=1-(PctCov/100)) %>%
  group_by(SiteID, PYear, PMonth, Transect) %>% 
  #creates a column to indicate if in that particular PLOT, there is more than one strata, meaning 'ground' and 'tree' layers. If so, a 'Y' is added, even if a single species was sampled in the tree layer
  mutate(RecordedStrata = ifelse(sum(code)>=1, "Y", "N")) %>% 
  select(-code) %>% 
  ungroup() %>% 
  group_by(SiteID, PYear, PMonth, Transect, OriginalSpeciesName) %>% 
  mutate(PctCov_100=(1-(prod(Step1)))*100) %>% 
  summarise(PctCov = sum(PctCov),
            SampledAreaOriginal = max(SampledAreaOriginal),
            RecordedStrata = dplyr::first(RecordedStrata),
            PctCov_100=dplyr::first(PctCov_100),
            NumberOfStrata = n_distinct(layer)) %>% 
  ungroup()

forestfinal %>% summarize(min = min(PctCov, na.rm = TRUE),
            mean = mean(PctCov, na.rm = TRUE),
            max = max(PctCov, na.rm = TRUE),
            min_100 = min(PctCov_100, na.rm = TRUE),
            mean_100 = mean(PctCov_100, na.rm = TRUE),
            max_100 = max(PctCov_100, na.rm = TRUE))

```

Joining all ecosystems together

```{r}

IL_CTAP <- forestfinal %>% 
  rbind(grassfinal) %>% 
  rbind(wetfinal) %>% 
  mutate(Dataset = "IL_CTAP") %>% 
  rename(SampledArea = SampledAreaOriginal) %>% 
  #
  # group_by(Dataset, SiteID, Transect, PYear, OriginalSpeciesName, RecordedStrata) %>% 
  # # this is the column I need to sum because trees present in different strata had distinct area sampled, so they are different observations in the data.
  # dplyr::summarise(PctCov = sum(PctCov),
  #                  # different trees were sampled in the different sampled area, I am keeping the first one
  #                  SampledArea = dplyr::first(SampledArea)) %>% 
  # ungroup() %>% 
  # distinct() %>% 
  #
  select(Dataset, everything())

## checking for duplicates of multiple months per year of data collection (fixed above, by ecosystem)

# IL_CTAPMultipleMonth<-
IL_CTAP %>% 
  select(SiteID, Transect, PYear, PMonth) %>% 
  distinct() %>% 
  group_by(SiteID, Transect, PYear) %>% 
  summarize(n_obs = n()) %>% 
  filter(n_obs > 1) %>% # None!
  ungroup()
# write.csv(IL_CTAPMultipleMonth, "/home/shares/neon-inv/data_paper/code_by_dataset/extra_csv_CTAP/IL_CTAPMultipleMonth.csv")

# IL_CTAPtest <- IL_CTAP %>% 
#   group_by(SiteID, PYear, Transect) %>% 
#   count(PMonth) %>% ungroup() # count the obaservations per site, year, transect and month
# 
# MonthsKeep <- IL_CTAPMultipleMonth %>% 
#   left_join(IL_CTAPtest) %>% # add information on the number of observations
#   group_by(SiteID, PYear, Transect) %>% 
#   top_n(1, n) %>% # filter the months with larger number of observations to be the ones we keep
#   group_by(SiteID, PYear, Transect) %>% 
#   slice(which.max(PMonth)) %>%  # there was one instance that a month had the excat same number of observations, so keeping the latest month of data collection
#   select(-n_obs, -n)
# 
# # fixing the database
# IL_CTAP <- IL_CTAP %>% 
#   left_join(MonthsKeep, by = c("SiteID", "PYear", "Transect")) %>% 
#   mutate(PMonth = ifelse(is.na(PMonth.y), PMonth.x, PMonth.y)) %>% 
#   select(-PMonth.y, -PMonth.x)


## checking for duplicates per species per plot per year
IL_CTAP %>% 
  group_by(SiteID, Transect, OriginalSpeciesName, PYear) %>% 
  summarise(n_obs = n()) %>% 
  filter(n_obs > 1) #None!
```

Matching species names
```{r}
tax <- read_csv(file.path(shared_data_dir, "code_by_dataset/taxonomy/taxonomy_temp10_revised.csv"))
tax <- tax %>%
  select(Accepted.Symbol, Scientific.Name, bestname, inv_L48) %>% 
  distinct() %>% 
  filter(!is.na(Accepted.Symbol))
glimpse(tax)

sppcodes <- read_csv("/home/shares/neon-inv/output_files/USDA_Plants_ScientificNames.csv")
head(sppcodes) 
# sppcodes <- sppcodes %>% select(-Synonym.Symbol)

sppfixed <- read_csv(file.path(data_extra, "IL_CTAPsppFixed.csv"))

IL_CTAPspp <- data.frame(unique(IL_CTAP$OriginalSpeciesName))
IL_CTAPspp <- IL_CTAPspp %>%
  rename(OriginalSpeciesName = 1) %>%
  separate(OriginalSpeciesName, c("Genus", "Species", "Extra"), remove = F) %>% 
  mutate(Species = ifelse(Species == "sp", "", 
                   ifelse(is.na(Species), "", Species)),
         Scientific.Name = paste0(Genus, " ", Species),
         Scientific.Name = ifelse(Scientific.Name == "NA NA", "", Scientific.Name),
         #removes any space after the end of any column containing characters
         across(where(is.character), str_trim)) %>% 
  select(-Genus, -Species, -Extra) 

## fixing typos on species names
IL_CTAPspp$Scientific.Name <- gsub("Amphicarpa bracteata", "Amphicarpaea bracteata", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Arisaema tryphyllum", "Arisaema triphyllum", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Carex pennsylvanicus", "Carex pensylvanica", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Carya illinoensis", "Carya illinoinensis", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Desmodium canadensis", "Desmodium canadense", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Echinochloa crusgalli", "Echinochloa crus-galli", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Eupatorium rugusom", "Eupatorium rugosum", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Hibiscus lasiocarpus", "Hibiscus lasiocarpos", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Lysimachia numeralia", "Lysimachia nummularia", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Quercus muhlenbergii", "Quercus muehlenbergii", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Rubus pensylvanicus", "Rubus pensilvanicus", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Samolus valerandii", "Samolus valerandi", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Smilax lasioneuron", "Smilax lasioneura", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Viburnum rafinesquianum", "Viburnum rafinesqueanum", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Viola sororis", "Viola sororia", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Acer saccarum", "Acer saccharum", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Lonicera mackii", "Lonicera maackii", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Gleditsia tricanthos", "Gleditsia triacanthos", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Maclura pomfera", "Maclura pomifera", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Wolffia braziliensis", "Wolffia brasiliensis", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Scirpus tabernaemontanii", "Scirpus tabernaemontani", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Polygonum sagitatum", "Polygonum sagittatum", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Polygonum sogitatum", "Polygonum sagittatum", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Plantago ruelli", "Plantago rugelii", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Nuphar luteum", "Nuphar lutea", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Carex strica", "Carex stricta", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Carex ovales", "Carex ovalis", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Aster symplex", "Aster simplex", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Aster lateriflora", "Aster lateriflorus", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Trifolium pretense", "Trifolium pratense", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Trifolium reptans", "Trifolium repens", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Tridens flavis", "Tridens flavus", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Solidago memoralis", "Solidago nemoralis", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Setaria pumil", "Setaria pumila", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Plantago lugelii", "Plantago rugelii", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Opuntia humifisa", "Opuntia humifusa", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Liparis lilifolia", "Liparis liliifolia", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Helianthus maximilianii", "Helianthus maximiliani", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Erysimum hieracifolium", "Erysimum hieraciifolium", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Cyperus grayioides", "Cyperus grayoides", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Cirsium canadense", "Cirsium canescens", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Robinia NA", "Robinia", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Acer rubra", "Acer rubrum", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Achillea milliflora", "Achillea multiflora", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Carex muhlenbergii", "Carex muehlenbergii", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Setaria pumilaa", "Setaria pumila", IL_CTAPspp$Scientific.Name)
IL_CTAPspp$Scientific.Name <- gsub("Aster punicens", "Aster puniceus", IL_CTAPspp$Scientific.Name)

IL_CTAPspp <- IL_CTAPspp %>%
#dataframe below fixes scientific names to match Ian's taxonomy
  left_join(select(sppfixed, OriginalSpeciesName, Scientific.Name), by = "OriginalSpeciesName") %>% 
  mutate(Scientific.Name = ifelse(is.na(Scientific.Name.y), Scientific.Name.x, Scientific.Name.y)) %>% 
  select(-Scientific.Name.x, -Scientific.Name.y) %>% 
  # I might need to add the IL_CTAPspp2Fixed.csv' here, before Ian
  # left join Ian's taxonomy
  left_join(tax, by = "Scientific.Name") %>%
  mutate(Accepted.Symbol = ifelse(Accepted.Symbol=="7-Feb", "FEBR7", Accepted.Symbol)) %>% 
  group_by(OriginalSpeciesName) %>% 
  add_tally() %>% 
  filter(!(is.na(inv_L48) & n==2),
         is.na(Accepted.Symbol)|Accepted.Symbol != "LEFR5") %>% 
  select(-n) 

#Synonym.Symbol
codesremove <- c("ASSA13", "CRCO18", "PALA30", "PAMI9", "OERH2", "ASVI5", "SCAM2", "WOPU")

tax2 <- tax %>% select(Accepted.Symbol, bestname, inv_L48) %>% distinct() 
tax3 <- tax %>% select(Scientific.Name, inv_L48) %>% distinct() 


IL_CTAPspp2 <- IL_CTAPspp %>%  filter(is.na(Accepted.Symbol) & is.na(inv_L48)) %>% 
  left_join(sppcodes, by = "Scientific.Name") %>%
  distinct(Accepted.Symbol.y, .keep_all = T) %>% 
  filter(!Synonym.Symbol %in% codesremove) %>% 
  mutate(Accepted.Symbol = ifelse(is.na(Accepted.Symbol.x), Accepted.Symbol.y, Accepted.Symbol.x)) %>% 
  select(-Accepted.Symbol.x, -Accepted.Symbol.y, -Synonym.Symbol) %>% 
  left_join(filter(sppcodes, is.na(Synonym.Symbol)) , by = "Accepted.Symbol") %>% 
  mutate(bestname = ifelse(is.na(bestname), Scientific.Name.y, bestname)) %>% 
  rename(Scientific.Name = Scientific.Name.x) %>% 
  select(-Scientific.Name.y, -Synonym.Symbol) %>% 
  mutate(bestname = sub(" var.*", "", bestname),
         bestname = sub(" ssp.*", "", bestname),
         #removes any space after the end of any column containing characters
         across(where(is.character), str_trim)) %>% 
    # left join Ian's taxonomy
  left_join(tax2, by = c("bestname", "Accepted.Symbol")) %>% 
  mutate(inv_L48 = ifelse(is.na(inv_L48.x), inv_L48.y, inv_L48.x)) %>% 
  select(-inv_L48.x, -inv_L48.y) %>% 
  left_join(tax3, by = c("bestname" = "Scientific.Name")) %>% 
  mutate(inv_L48 = ifelse(is.na(inv_L48.x), inv_L48.y, inv_L48.x)) %>% 
  select(-inv_L48.x, -inv_L48.y)

sppfixed2 <- read_csv(file.path(data_extra, "IL_CTAPspp3Fixed.csv"))
sppfixed3 <- read_csv(file.path(data_extra, "IL_CTAPspp4Fixed.csv"))

IL_CTAPsppfinal <- IL_CTAPspp %>% 
  filter(!(is.na(Accepted.Symbol) & is.na(inv_L48))) %>% 
  rbind(IL_CTAPspp2) %>% 
  left_join(sppfixed2, by = c("OriginalSpeciesName", "Scientific.Name", "bestname", "Accepted.Symbol")) %>% 
  mutate(inv_L48 = ifelse(is.na(inv_L48.x), inv_L48.y, inv_L48.x)) %>% 
  select(-inv_L48.x, -inv_L48.y) %>% 
  left_join(sppfixed3, by = c("OriginalSpeciesName", "Scientific.Name", "bestname", "Accepted.Symbol")) %>% 
  mutate(inv_L48 = ifelse(is.na(inv_L48.x), inv_L48.y, inv_L48.x)) %>% 
  select(-inv_L48.x, -inv_L48.y)

```

Joining the fixed species table to the main database
```{r}
IL_CTAP2 <- left_join(IL_CTAP, IL_CTAPsppfinal, by = "OriginalSpeciesName")
glimpse(IL_CTAP2)
is.na(IL_CTAP2) <- IL_CTAP2==''

IL_CTAP2 %>% group_by(inv_L48) %>% tally()

```

Including/fixing Exotic status hand correct 
```{r}
hand_exo <- read_csv(file.path(shared_data_dir, "code_by_dataset/taxonomy/multStatusSpL48LP.csv"))
head(hand_exo)

hand_exo <- hand_exo %>%
  rename(Accepted.Symbol = SpCode) %>%
  select(Accepted.Symbol, 'FINAL DECISION (L48)') %>%
  distinct()

IL_CTAP2 <- IL_CTAP2 %>% left_join(hand_exo)
IL_CTAP2 <- IL_CTAP2 %>% mutate(inv_L48 = ifelse(is.na(`FINAL DECISION (L48)`), inv_L48, `FINAL DECISION (L48)`)) %>% select(-`FINAL DECISION (L48)`)

#check one that was NI and should now be N
IL_CTAP2 %>% filter(Accepted.Symbol=="AMBRO")

```


Adding the coordinates
```{r}
latlong <- site_info %>% 
  select(SiteID, Lat, Lon) %>% 
  rename(Long = Lon) %>% distinct()

IL_CTAP2 <- IL_CTAP2 %>% 
  left_join(latlong, by = "SiteID") %>% 
  #creates unique plotIDs
  mutate(Plot = paste0(SiteID, "_", Transect))

glimpse(IL_CTAP2)
```

Fixing coordinates for site 000501G according to David Zaya instructions
```{r}
IL_CTAP2 <- IL_CTAP2 %>% 
  mutate(Lat = ifelse(SiteID=="000501G", 42.47399, Lat),
         Long = ifelse(SiteID=="000501G", -90.18289, Long))
```

Defining PlotArea
```{r}
PlotAreaW <- wtree %>% select(PSID, SiteID, PYear, Squmetr) %>% filter(PSID=="None") %>% rename(Squrmetrs=Squmetr) %>% distinct()
PlotAreaG <- gtree %>% select(PSID, SiteID, PYear, Squrmetrs) %>% filter(PSID=="None") %>% distinct()

PlotArea <- PlotAreaW %>% rbind(PlotAreaG) %>% select(-PSID) %>% rename(PlotArea.m2 = Squrmetrs)

```


After standardizing taxonomy there are four instances in which species with the same accepted name are duplicated.
```{r}
# are there multiple species code per plot? Yes
IL_CTAP2 %>% 
  group_by(Dataset, Plot, Accepted.Symbol, PYear) %>% 
  filter(!is.na(Accepted.Symbol)) %>% 
  summarise(n_obs = n()) %>% 
  filter(n_obs > 1)

#fixing it
IL_CTAP2 <- IL_CTAP2 %>% 
  ## check if all columns are listed here
  group_by(Dataset, SiteID, Transect, PYear, PMonth, 
           RecordedStrata,
           Accepted.Symbol, inv_L48, Lat, Long, Plot) %>% 
  # this is the column I need to sum because trees present in different strata had distinct area sampled, so they are different observations in the data.
  dplyr::summarise(PctCov = sum(PctCov),
                   PctCov_100 = sum(PctCov_100),
                   # different trees were sampled in the different sampled area, I am keeping the first one
                   SampledArea = max(SampledArea),
                   OriginalSpeciesName = dplyr::first(OriginalSpeciesName),
                   # the tax file I am using here is returning multiple Scientific or bestname for the same species code. This is fixed when dataset is incorporated to the Database
                   Scientific.Name = dplyr::first(Scientific.Name),
                   bestname = dplyr::first(bestname)) %>% 
  ungroup() %>% 
  distinct() %>% 
  left_join(PlotArea) %>% mutate(PlotArea.m2=ifelse(grepl("F$", SiteID), 500,
                                                    ifelse(is.na(PlotArea.m2), 2050, PlotArea.m2))) %>% 
  select(Dataset, everything())

length(colnames(IL_CTAP2))

## checking for duplicates per species per plot per year
IL_CTAP2 %>% 
  group_by(Dataset, Plot, Accepted.Symbol, PYear) %>% 
  filter(!is.na(Accepted.Symbol)) %>% 
  summarise(n_obs = n()) %>% 
  filter(n_obs > 1)
```

Checking range of PctCov
```{r}
IL_CTAP2 %>%
  summarize(min = min(PctCov, na.rm = TRUE),
            mean = mean(PctCov, na.rm = TRUE),
            max = max(PctCov, na.rm = TRUE),
            min_100 = min(PctCov_100, na.rm = TRUE),
            mean_100 = mean(PctCov_100, na.rm = TRUE),
            max_100 = max(PctCov_100, na.rm = TRUE))
```


Exporting final file
```{r}
IL_CTAP2 %>% write.csv(file.path(data_output, "IL_CTAP_DataPaper09132022.csv"), row.names = F)
```

