---
title: "NWCA_code_datapaper"
output: html_document
editor_options: 
  chunk_output_type: console
notes: the raw file of 2011 data was given to Laís from Ian Pearse via email. The 2016 data Laís downloaded from the website and included in the database. https://www.epa.gov/national-aquatic-resource-surveys/data-national-aquatic-resource-surveys
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages
```{r}
library(tidyverse)
library(lubridate)
```

Paths to data
```{r}
shared_data_dir <- "/home/shares/neon-inv/data_paper"
data_raw <- file.path(shared_data_dir, "data_by_dataset/archived")
data_output <- file.path(shared_data_dir,"data_by_dataset")

```

Importing dataset 2011
```{r}
NWCA2011 <- read_xlsx(file.path(data_raw, "nwca11_plant_wide.xlsx"), sheet = "nwca11_plant_wide")
glimpse(NWCA2011)

NWCA2011 <- NWCA2011 %>%
  rename(Year = DSGN_CYCLE) %>%
  filter(!is.na(COVER)) %>% 
  #unique id matches sites across years
  #plots were not exactly the same location within site across years
  select(UNIQUE_ID, SITE_ID, PLOT, DATE_COL, Year, COVER, SPECIES) %>%
  rename(OriginalName = SPECIES) %>%
  #code below fixes species (same SPECIES name) that are recorded more than once in a plot
  #it seems that the updated file Gregg sent us has this issue fixed
  group_by(UNIQUE_ID, PLOT, DATE_COL, Year, SITE_ID, OriginalName) %>% 
  summarise(COVER = sum(COVER, na.rm=TRUE)) %>% 
  ungroup()
```

Importing dataset 2016
```{r}
NWCA2016 <- read_xlsx(file.path(data_raw, "nwca16_plant_wide.xlsx"), sheet = "nwca16_plant_wide")
glimpse(NWCA2016)

NWCA2016 <- NWCA2016 %>%
  rename(Year = DSGN_CYCLE) %>%
  filter(!is.na(COVER)) %>% 
  #unique id matches sites across years
  #plots were not exactly the same location within site across years
  select(UNIQUE_ID, SITE_ID, PLOT, DATE_COL, Year, COVER, SPECIES) %>%
  rename(OriginalName = SPECIES) %>%
  #code below fixes species (same SPECIES name) that are recorded more than once in a plot
  #it seems that the updated file Gregg sent us has this issue fixed
  group_by(UNIQUE_ID, PLOT, DATE_COL, Year, SITE_ID, OriginalName) %>% 
  summarise(COVER = sum(COVER, na.rm=TRUE)) %>% 
  ungroup()

```

Importing information on site lat/long
```{r}
latlong2011 <- read_csv(file.path(shared_data_dir, "code_by_dataset/extra_csv_NWCA/nwca2011_siteinfo.csv"))
glimpse(latlong2011)
latlong2011 <- latlong2011 %>% select(SITE_ID, AA_CENTER_LAT, AA_CENTER_LON) %>%
  rename(Lat = AA_CENTER_LAT,
         Long = AA_CENTER_LON) %>% 
  distinct(SITE_ID, .keep_all = T)

latlong2016 <- read_csv(file.path(shared_data_dir, "code_by_dataset/extra_csv_NWCA/nwca_2016_site_information_-_data_csv.csv"))
glimpse(latlong2016)
latlong2016 <- latlong2016 %>% select(SITE_ID, LAT_DD83, LON_DD83) %>%
  rename(Lat = LAT_DD83,
         Long = LON_DD83)  %>% 
  distinct(SITE_ID, .keep_all = T)
```

Finding resampled sites and giving them unique IDs
```{r}
#finding sites that were resampled in 2016, and giving them a single site ID across years
#plots were likely not resurveyed in the exact same location in between the two campaigns

#getting distinct UNIQUE_ID per Site_ID
UniqueIDs <- NWCA2011 %>% 
  select(UNIQUE_ID) %>% 
  rbind(select(NWCA2016, UNIQUE_ID)) %>% 
  distinct()  
  

ResampledSites <- read_xlsx(file.path(data_raw, "nars_allsurvey_site_id_crosswalk-wide-and-long-250122.xlsx"), sheet = "WIDE_NARS_ALLSurvey_SITE_ID CRO") %>% 
  filter(UNIQUE_ID %in% UniqueIDs$UNIQUE_ID) %>% 
  select(UNIQUE_ID, `2011`, `2016`) %>% 
  mutate(code = ifelse(`2011` == "NA" | `2016` == "NA", 0, 1),
         stateID = str_sub(UNIQUE_ID, 5, -7)) %>% 
  filter(code==1) %>% 
  group_by(stateID, UNIQUE_ID) %>% 
  mutate(UniqueSiteID = paste0("NWCA_",stateID, "_", cur_group_id())) %>% 
  ungroup() %>% 
  rename(SiteID2011 = `2011`,
         SiteID2016 = `2016`) %>% 
  select(-stateID)
  
```

Adding lat/long info
```{r}
NWCA2011 <- left_join(NWCA2011, latlong2011, by = "SITE_ID")
glimpse(NWCA2011)

NWCA2016 <- left_join(NWCA2016, latlong2016, by = c("SITE_ID"))
glimpse(NWCA2016)
```

Adding new site ID information
```{r}
NWCA2011 <- NWCA2011 %>% 
  left_join(select(ResampledSites, c(SiteID2011, UniqueSiteID)), by = c("SITE_ID"="SiteID2011")) %>% 
  rename(Original.Site = SITE_ID) %>% 
  mutate(UniqueSiteID = ifelse(is.na(UniqueSiteID), Original.Site, UniqueSiteID)) %>% 
  select(-UNIQUE_ID)
glimpse(NWCA2011)

NWCA2016 <- NWCA2016 %>% 
  left_join(select(ResampledSites, c(SiteID2016, UniqueSiteID)), by = c("SITE_ID"="SiteID2016")) %>% 
  rename(Original.Site = SITE_ID) %>% 
  mutate(UniqueSiteID = ifelse(is.na(UniqueSiteID), Original.Site, UniqueSiteID)) %>%
  select(-UNIQUE_ID)
glimpse(NWCA2016)
```


Merging the two datasets
```{r}

NWCA <- bind_rows(NWCA2011, NWCA2016) 
NWCA <- NWCA %>% 
  group_by(Original.Site, PLOT) %>%
  mutate(UniquePlotID = paste0("NWCA_", cur_group_id()),
         #https://www.epa.gov/sites/default/files/2013-11/documents/fom_with_errata.pdf
         PlotArea_m2 = 100) %>% 
  #removes true absences as no other dataset has them.
  filter(COVER > 0)
glimpse(NWCA)

```

Matching species names
```{r}
tax <- read_csv(file.path(shared_data_dir, "code_by_dataset/taxonomy/taxonomy_temp10_revised.csv"))
tax <- tax %>%
  select(Accepted.Symbol, Scientific.Name, bestname, inv_L48) %>%
  filter(!is.na(inv_L48))
glimpse(tax)

sppcodes <- read_csv("/home/shares/neon-inv/output_files/USDA_Plants_ScientificNames.csv")
head(sppcodes) 
sppcodes <- sppcodes %>% select(-Synonym.Symbol)

NWCAspp <- data.frame(unique(NWCA$OriginalName))
NWCAspp <- NWCAspp %>%
  rename(OriginalName = 1) %>%
  # mutate(OriginalName = ifelse(OriginalName=="ACER xFREEMANII", "Acer freemanii", OriginalName)) %>%
  mutate(Scientific.Name = str_to_lower(OriginalName),
         Scientific.Name = str_to_sentence(Scientific.Name)) %>%
  left_join(tax) %>%
  mutate(Accepted.Symbol = ifelse(Accepted.Symbol=="7-Feb", "FEBR7", Accepted.Symbol)) %>%
  left_join(sppcodes, by = "Scientific.Name") %>%
  mutate(Accepted.Symbol = ifelse(is.na(Accepted.Symbol.x), Accepted.Symbol.y, Accepted.Symbol.x)) %>%
  select(-Accepted.Symbol.x, -Accepted.Symbol.y) %>%
  distinct(OriginalName, .keep_all = T)

```

Joining the fixed species table to the main database
```{r}
NWCA2 <- left_join(NWCA, NWCAspp)
glimpse(NWCA2)
is.na(NWCA2) <- NWCA2=='' ## ISSUE - this seems to not be working

NWCA2 %>% group_by(inv_L48) %>% tally()
```

Including/fixing Exotic status hand correct 
```{r}
hand_exo <- read_csv(file.path(shared_data_dir, "code_by_dataset/taxonomy/multStatusSpL48.csv"))
head(hand_exo)

hand_exo <- hand_exo %>%
  rename(Accepted.Symbol = SpCode) %>%
  select(Accepted.Symbol, 'FINAL DECISION (L48)') %>%
  distinct

NWCA2 <- NWCA2 %>% left_join(hand_exo)
NWCA2 <- NWCA2 %>% mutate(inv_L48 = ifelse(is.na(`FINAL DECISION (L48)`), inv_L48, `FINAL DECISION (L48)`)) %>% select(-`FINAL DECISION (L48)`)

#check one that was NA and should now be I
NWCA2 %>% filter(Accepted.Symbol=="BRIN2") %>% select(Accepted.Symbol, COVER, inv_L48)

glimpse(NWCA2)

```

Data checks
```{r}
#how many plots have no coordinate info
NWCA2 %>% 
  select(UniquePlotID, Lat, Long) %>% 
  filter(is.na(Lat) | is.na(Long)) %>% 
  count() ## NONE!

```


Exporting final dataset
```{r}
write.csv(NWCA2, file.path(data_output, "NWCA2011-2016_04202022.csv"))
```

