---
title: "NWCA_code_datapaper"
output: html_document
editor_options: 
  chunk_output_type: console
notes: the raw file of 2011 data was given to Laís from Ian Pearse via email. The 2016 data Laís downloaded from the website and included in the database. https://www.epa.gov/national-aquatic-resource-surveys/data-national-aquatic-resource-surveys
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages
```{r}
library(tidyverse)
library(lubridate)
```

Paths to data
```{r}
shared_data_dir <- "/home/shares/neon-inv/data_paper"
data_raw <- file.path(shared_data_dir, "data_by_dataset/archived")
data_output <- file.path(shared_data_dir,"data_by_dataset")

```

Importing dataset 2011
```{r}
NWCA2011 <- read_xlsx(file.path(data_raw, "nwca11_plant_wide.xlsx"), sheet = "nwca11_plant_wide")
glimpse(NWCA2011)

# checking why aggregating the plots per UID might not be correct
NWCA2011 %>% distinct(UID) %>% count()
NWCA2011 %>% distinct(SITE_ID) %>% count()
NWCA2011 %>% group_by(UID, PLOT) %>% count() %>% nrow()
NWCA2011 %>% group_by(SITE_ID, PLOT) %>% count() %>% nrow()

# the issue is that SITE_ID and UID have different number of categories. For the previous data, I was giving the plots new names based on UID & PLOT, but I think it should have been SITE_ID & PLOT
# UID refers to a unique id per visit per plot per site. Some sites were visited more than once within a year.
# For plots with "VISIT_NO" of 1 and 2, I am keeping number 2 only (the latest visit)
# this will affect the total number of plots, total number of observations and the summaries of the very final database!!!!!!

# previously, I was finding the latest revisit to a plot in a year. But, after talking to Gregg from EPA, he said that the first visit of the year is considered by them as the official one. So, I am using the same.

NWCA2011 <- NWCA2011 %>%
  filter(VISIT_NO == 1) %>% 
  rename(Year = DSGN_CYCLE) %>%
  filter(!is.na(COVER)) %>% 
  #unique id matches sites across years
  #plots were not exactly the same location within site across years
  select(UNIQUE_ID, SITE_ID, PLOT, DATE_COL, Year, COVER, SPECIES, VISIT_NO) %>%
  rename(OriginalName = SPECIES) %>%
  #code below fixes species (same SPECIES name) that are recorded more than once in a plot
  #it seems that the updated file Gregg sent us has this issue fixed
  group_by(UNIQUE_ID, PLOT, DATE_COL, Year, SITE_ID, OriginalName, VISIT_NO) %>% 
  summarise(COVER = sum(COVER, na.rm=TRUE)) %>% 
  ungroup()
```

Importing dataset 2016
```{r}
NWCA2016 <- read_xlsx(file.path(data_raw, "nwca16_plant_wide.xlsx"), sheet = "nwca16_plant_wide")
glimpse(NWCA2016)

# previously, I was finding the latest revisit to a plot in a year. But, after talking to Gregg from EPA, he said that the first visit of the year is considered by them as the official one. So, I am using the same.

NWCA2016 <- NWCA2016 %>%
  filter(VISIT_NO == 1) %>% 
  rename(Year = DSGN_CYCLE) %>%
  filter(!is.na(COVER)) %>% 
  #unique id matches sites across years
  #plots were not exactly the same location within site across years
  select(UNIQUE_ID, SITE_ID, PLOT, DATE_COL, Year, COVER, SPECIES, VISIT_NO) %>%
  rename(OriginalName = SPECIES) %>%
  #code below fixes species (same SPECIES name) that are recorded more than once in a plot
  #it seems that the updated file Gregg sent us has this issue fixed
  group_by(UNIQUE_ID, PLOT, DATE_COL, Year, SITE_ID, OriginalName, VISIT_NO) %>% 
  summarise(COVER = sum(COVER, na.rm=TRUE)) %>% 
  ungroup()

```

Importing information on site lat/long
```{r}
latlong2011 <- read_csv(file.path(shared_data_dir, "code_by_dataset/extra_csv_NWCA/nwca2011_siteinfo.csv"))
glimpse(latlong2011)
latlong2011 <- latlong2011 %>% select(SITE_ID, VISIT_NO, AA_CENTER_LAT, AA_CENTER_LON) %>%
  filter(VISIT_NO == 1) %>% 
  rename(Lat = AA_CENTER_LAT,
         Long = AA_CENTER_LON) %>% 
  distinct(SITE_ID, .keep_all = T)

latlong2016 <- read_xlsx(file.path(shared_data_dir, "code_by_dataset/extra_csv_NWCA/nwca16_siteinfo.xlsx"), sheet = "nwca16_siteinfo")
glimpse(latlong2016)
latlong2016 <- latlong2016 %>% select(SITE_ID, VISIT_NO, LAT_DD83, LON_DD83) %>%
  filter(VISIT_NO == 1) %>% 
  rename(Lat = LAT_DD83,
         Long = LON_DD83)  %>% 
  distinct(SITE_ID, .keep_all = T)
```

Finding resampled sites and giving them unique IDs
```{r}
#finding sites that were resampled in 2016, and giving them a single site ID across years
#plots were likely not resurveyed in the exact same location in between the two campaigns

#getting distinct UNIQUE_ID per Site_ID
UniqueIDs <- NWCA2011 %>% 
  select(UNIQUE_ID) %>% 
  rbind(select(NWCA2016, UNIQUE_ID)) %>% 
  distinct()  
  

ResampledSites <- read_xlsx(file.path(data_raw, "nars_allsurvey_site_id_crosswalk-wide-and-long-250122.xlsx"), sheet = "WIDE_NARS_ALLSurvey_SITE_ID CRO") %>% 
  filter(UNIQUE_ID %in% UniqueIDs$UNIQUE_ID) %>% 
  select(UNIQUE_ID, `2011`, `2016`) %>% 
  mutate(code = ifelse(`2011` == "NA" | `2016` == "NA", 0, 1),
         stateID = str_sub(UNIQUE_ID, 5, -7)) %>% 
  filter(code==1) %>% 
  group_by(stateID, UNIQUE_ID) %>% 
  mutate(UniqueSiteID = paste0("NWCA_",stateID, "_", cur_group_id())) %>% 
  ungroup() %>% 
  rename(SiteID2011 = `2011`,
         SiteID2016 = `2016`) %>% 
  select(-stateID)
  
```

Adding lat/long info
```{r}
NWCA2011 <- left_join(NWCA2011, latlong2011, by = c("SITE_ID", "VISIT_NO"))
glimpse(NWCA2011)

NWCA2016 <- left_join(NWCA2016, latlong2016, by = c("SITE_ID", "VISIT_NO"))
glimpse(NWCA2016)
```

Adding new site ID information
```{r}
NWCA2011 <- NWCA2011 %>% 
  left_join(select(ResampledSites, c(SiteID2011, UniqueSiteID)), by = c("SITE_ID"="SiteID2011")) %>% 
  rename(Original.Site = SITE_ID) %>% 
  mutate(UniqueSiteID = ifelse(is.na(UniqueSiteID), Original.Site, UniqueSiteID)) %>% 
  select(-UNIQUE_ID)
glimpse(NWCA2011)

NWCA2016 <- NWCA2016 %>% 
  left_join(select(ResampledSites, c(SiteID2016, UniqueSiteID)), by = c("SITE_ID"="SiteID2016")) %>% 
  rename(Original.Site = SITE_ID) %>% 
  mutate(UniqueSiteID = ifelse(is.na(UniqueSiteID), Original.Site, UniqueSiteID)) %>%
  select(-UNIQUE_ID)
glimpse(NWCA2016)
```


Merging the two datasets
```{r}

NWCA <- bind_rows(NWCA2011, NWCA2016) 

NWCA %>% distinct(Original.Site) %>% count()
NWCA2011_2016_02222022 %>% distinct(SITE_ID) %>% count()

glimpse(NWCA)
NWCA <- NWCA %>% 
  group_by(Original.Site, PLOT) %>%
  mutate(UniquePlotID = paste0("NWCA_", cur_group_id()),
         #https://www.epa.gov/sites/default/files/2013-11/documents/fom_with_errata.pdf
         PlotArea_m2 = 100) %>% 
  ungroup() %>% 
  #removes true absences as no other dataset has them.
  filter(COVER > 0)

NWCA %>% distinct(UniquePlotID) %>%  count()

glimpse(NWCA)

```

Matching species names
```{r}
tax <- read_csv(file.path(shared_data_dir, "code_by_dataset/taxonomy/taxonomy_temp10_revised.csv"))
tax <- tax %>%
  select(Accepted.Symbol, Scientific.Name, bestname, inv_L48) %>%
  filter(!is.na(inv_L48))
glimpse(tax)

sppcodes <- read_csv("/home/shares/neon-inv/output_files/USDA_Plants_ScientificNames.csv")
head(sppcodes) 
sppcodes <- sppcodes %>% select(-Synonym.Symbol)

NWCAspp <- data.frame(unique(NWCA$OriginalName))
NWCAspp <- NWCAspp %>%
  rename(OriginalName = 1) %>%
  # mutate(OriginalName = ifelse(OriginalName=="ACER xFREEMANII", "Acer freemanii", OriginalName)) %>%
  mutate(Scientific.Name = str_to_lower(OriginalName),
         Scientific.Name = str_to_sentence(Scientific.Name)) %>%
  left_join(tax) %>%
  mutate(Accepted.Symbol = ifelse(Accepted.Symbol=="7-Feb", "FEBR7", Accepted.Symbol)) %>%
  left_join(sppcodes, by = "Scientific.Name") %>%
  mutate(Accepted.Symbol = ifelse(is.na(Accepted.Symbol.x), Accepted.Symbol.y, Accepted.Symbol.x)) %>%
  select(-Accepted.Symbol.x, -Accepted.Symbol.y) %>%
  distinct(OriginalName, .keep_all = T)

```

Joining the fixed species table to the main database
```{r}
NWCA2 <- left_join(NWCA, NWCAspp)
glimpse(NWCA2)
is.na(NWCA2) <- NWCA2=='' ## ISSUE - this seems to not be working

NWCA2 %>% group_by(inv_L48) %>% tally()
```

Including/fixing Exotic status hand correct 
```{r}
hand_exo <- read_csv(file.path(shared_data_dir, "code_by_dataset/taxonomy/multStatusSpL48.csv"))
head(hand_exo)

hand_exo <- hand_exo %>%
  rename(Accepted.Symbol = SpCode) %>%
  select(Accepted.Symbol, 'FINAL DECISION (L48)') %>%
  distinct

NWCA2 <- NWCA2 %>% left_join(hand_exo)
NWCA2 <- NWCA2 %>% mutate(inv_L48 = ifelse(is.na(`FINAL DECISION (L48)`), inv_L48, `FINAL DECISION (L48)`)) %>% select(-`FINAL DECISION (L48)`)

#check one that was NA and should now be I
NWCA2 %>% filter(Accepted.Symbol=="BRIN2") %>% select(Accepted.Symbol, COVER, inv_L48)

glimpse(NWCA2)

```

Data checks
```{r}
#how many plots have no coordinate info
NWCA2 %>% 
  select(UniquePlotID, Lat, Long) %>% 
  filter(is.na(Lat) | is.na(Long)) %>% 
  count() ## NONE!

```

Comparing file before the updated NWCA raw data
```{r}
#most recent file prior to Gregg's new raw data
NWCA2011_2016_02222022 <- read_csv("data_by_dataset/archived/NWCA2011-2016_02222022.csv")
glimpse(NWCA2011_2016_02222022)

NWCA2011_2016_02222022 %>% 
  distinct(UniquePlotID) %>% 
  count() ##11845 plots

NWCA2011_2016_02222022 %>% 
  distinct(SITE_ID) %>% 
  count() ##2194 sites

NWCA2011_2016_02222022 %>% 
  count(PLOT) 
#    PLOT     n
#   <dbl> <int>
# 1     1 38739
# 2     2 38990
# 3     3 38060
# 4     4 37319
# 5     5 37336

NWCA2011_2016_02222022 %>% 
  group_by(Year) %>% 
  count(PLOT)
#     Year  PLOT     n
#    <dbl> <dbl> <int>
#  1  2011     1 18208
#  2  2011     2 18458
#  3  2011     3 17745
#  4  2011     4 17419
#  5  2011     5 17728
#  6  2016     1 20531
#  7  2016     2 20532
#  8  2016     3 20315
#  9  2016     4 19900
# 10  2016     5 19608

NWCA2 %>% 
  ungroup() %>% 
  distinct(Original.Site) %>% 
  count() ##2194 sites

NWCA2 %>% 
    ungroup() %>% 
  distinct(UniquePlotID) %>% 
  count() ##10896 plots

NWCA2 %>% 
  ungroup() %>% 
  count(PLOT)# not the same a before, only the latest survey plots sampled twice within a year were kept
#    PLOT     n
#   <dbl> <int>
# 1     1 35501
# 2     2 35529
# 3     3 34778
# 4     4 34101
# 5     5 34162

NWCA2 %>% 
  ungroup() %>% 
  group_by(Year) %>% 
  count(PLOT)
#     Year  PLOT     n
#    <dbl> <dbl> <int>
#  1  2011     1 16591
#  2  2011     2 16747
#  3  2011     3 16122
#  4  2011     4 15828
#  5  2011     5 16180
#  6  2016     1 18910
#  7  2016     2 18782
#  8  2016     3 18656
#  9  2016     4 18273
# 10  2016     5 17982
```


Exporting final dataset
```{r}
write.csv(NWCA2, file.path(data_output, "NWCA2011-2016_04212022.csv"))
```

